\chapter{Architecture Overview}
	Democritus' compiler is built off of Professor Stephen Edwards' \texttt{MicroC} compiler. 

\section{Compiler Overview}
  Several files make up the source code of the compiler. These include:
  \begin{itemize}
    \item \texttt{scanner.mll}: the OCamllex scanner.
    \item \texttt{ast.ml}: the abstract syntax tree, summarizing the overall structure of a Democritus program. 
    \item \texttt{parser.mly}: the Ocamlyacc parser. Tokens from the scanner are parsed into the abstract syntax tree in the parser.
    \item \texttt{semant.ml}: the semantic analyzer.
    \item \texttt{codegen.ml}: the LLVM IR code generator.
    \item \texttt{democritus.ml}: the overarching OCaml program that calls the four main steps of the compiler.
    \item \texttt{bindings.c}: a C file that provides facilitates low-level operations that interact with the OS through C functions, such as for threads, which is then compiled to LLVM bytecode.
  \end{itemize}
	\subsection{The Scanner}
    The scanner is simply a text scanner that parses text into various tokens, to then be interpreted by the parser. The regular expressions used by the scanner are listed in the language reference chapter.

	\subsection{The Parser}
    The parser is a token scanner that converts the tokens read into a valid abstract syntax tree of the program. If the program follows valid syntax, it will be parsed accordingly. Otherwise, compilation of code will yield a parse error. The structure of the program is as follows:

	\subsection{The Semantic Analyzer}

	\subsection{The Code Generator}
